The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2017-08-30 rnw/Main structure and modules of the Package.Rnw
update Andreas 29.08.2017(2)
update Andreas 30.08.2017(2)

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each point $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume density per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model. The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$. 
%In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase sampling can be applied with the \code{twophase()} function.

@@

The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (figure \ref{fig:concmphase_and_sae_a}). The basic assumption is that auxiliary information is available in two different frequencies, i.e. sample sizes. The phase $s_0$ provides a large number $n_0$ of auxiliary information, whereas the phase $s_1$ provides additional auxiliary information that is only available at $n_1$ subsamples of $s_0$. The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and $s_0$. The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$. From that it follows that we can define \textit{two nested} regression models. The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$. The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}. Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(0)}(x)$ in the so called \textit{reduced model} at the sample locations $x \in s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{0}(x)$. If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator. 
%In the \proglang{R} package \pkg{forestinventory}, estimators for three-phase sampling can be applied with the \code{threephase()} function.

@@

We enter the topic of small area estimation if we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (figure \ref{fig:concmphase_and_sae_b}). If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be to small to provide sufficient estimation accuracy. In this case, the application of multiphase estimation procedures can be efficient. However, the terrestrial sample size is too small to fit a separate regression model to them. The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ and then to apply this  model to the small area. A potential model bias within $G$ is then corrected for by using the model residuals derived in the small area $G$. If no terrestrial samples fall into a small area $G$ (i.e. $n_{2,G}=0$), this bias correction cannot be applied and the resulting \textit{synthetic} estimates are potentially biased. 
% Both two and three-phase small area estimators for synthetic and non-synthetic estimation can be applied by the \code{twophase()} and \code{threephase()} functions in the \proglang{R} package \pkg{forestinventory}.

@@

The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories. However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest). As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e. the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable. Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e. the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process. The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism. A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The randomly distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids due to minimization of travelling cost in the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point estimation as well as the variance estimation, which will in most cases by slightly overestimated \citep{mandallaz2013a}.

@@

\subsection{Package Structure}

In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}). The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline. The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global estimators}) or only for subregions of the entire inventory area (\textbf{small area estimators}). The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area. The first case allows for computing the exact mean $\bar{\pmb{Z}}$ of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$). In the second case, the mean of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$) can only be estimated by $\hat{\bar{\pmb{Z}}}$ which adds another component of uncertainty to the respective estimators. The package offers three small area estimators for both the exhaustive and non-exhaustive setup. The \textit{extended synthetic estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$. The resulting point estimates and variances are asymptotically design-unbiased. If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased. The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (\textit{one-phase estimation}) by the \code{onephase()} function. All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated samples.

@@

The following sections describe the mathematical details and the application of the multiphase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix. Special cases under cluster sampling are described in section \ref{sec:speccas_and_scen}.

>>>>

The two-phase or double-sampling estimators use inventory information from \textbf{two} nested samples which are commonly reffered to as \textit{phases} (figure \ref{fig:concmphase_and_sae_a}). The first phase $s_1$ comprises $n_1$ sample locations that each provide a set of explanatory variables described by the column vector $\pmb{Z}(x)\in{\Re^{p}}$ at each location $x \in s_1$. These explanatory variables are derived from auxiliary information that is available in high quantity within the forest area $F$. The second phase $s_2$ constitutes the terrestrial inventory conducted at $n_2$ subsamples of the large phase $s_1$ and provides the value of the target variable, i.e. the local density $Y(x)$ (e.g. the timber volume density per hectare). The set of explanatory variables at each sample location $x \in s_1$ is transformed into a prediction $\hat{Y}(x)$ of $Y(x)$ by the application of an oridinary least square (OLS) regression model. The basic idea of this setup is to boost the sample size by providing a large sample of less precise but cheap information about $Y(x)$ in $x \in s_1$, and to account for the lack of precision by collecting precise but costly information of $Y(x)$ at few sample locations $x \in s_2 \subset s_1$.

@@

The three-phase or triple-sampling estimators extend the principle of two-phase sampling and use inventory information from \textbf{three} nested samples (\textit{phases}) (figure \ref{fig:concmphase_and_sae_a}). The basic assumption is that auxiliary information is available in two different frequencies, i.e. sample sizes. The phase $s_0$ provides a large number $n_0$ of auxiliary information, whereas the phase $s_1$ provides additional auxiliary information that is only available at $n_1$ subsamples of $s_0$. The terrestrial information is then collected at a subsample $s_2$ of $s_1$ and thus $s_0$. The basic idea of three-phase sampling is that the additional set of auxiliary information $\pmb{Z}^{(1)}(x)\in{\Re^q}$ available in $s_1$ adds considerable explanatory power to the set of auxiliary information $\pmb{Z}^{(0)}(x)\in{\Re^p}$ available at all sample locations $x \in s_0$. From that it follows that we can define \textit{two nested} regression models. The \textit{full} set of predictor variables $\pmb{Z}^t(x)=(\pmb{Z}^{(0)t}(x),\pmb{Z}^{(1)t}(x))\in{\Re^{p+q}}$ in the sample $s_1$ can be used to calculate predictions $\hat{Y}_{1}(x)$ of $Y(x)$ at all sample locations $x \in s_1$. The regression model applicable to the $s_1$ phase is thus referred to as the \textit{full model}. Whereas we can only use the subset of auxiliary information $\pmb{Z}^{(0)}(x)$ in the so called \textit{reduced model} at the sample locations $x \in s_0$, it allows for creating a much larger sample of less accurate predictions $\hat{Y}_{0}(x)$. If the gain in model precision between the reduced and the full model as well as the sampling fraction between $s_0$ and $s_1$ is sufficiently large, the three-phase estimator normally leads to a further increase in estimation precision compared to the two-phase estimator.

@@

We enter the topic of small area estimation if we want to make an estimation for a subregion $G$ of the entire inventory area $F$ (figure \ref{fig:concmphase_and_sae_b}). If the spatial extent of the subregion is small with respect to the terrestrial sample density, the number of terrestrial sample locations within the subregion $n_{2,G}$ will be to small to provide sufficient estimation accuracy. In this case, the application of multi-phase estimation procedures can be particularly efficient. However, the terrestrial sample size in $G$ is often too small to fit a separate regression model. The idea is then to \textit{borrow strength} from the entire terrestrial sample $s_2$ of $F$ for model fitting, and then to apply this model in the small area. If $G$ provides terrestrial samples, a model bias in $G$ is corrected using the model residuals in $G$, leading to unbiased estimates. Otherwise, the resulting estimates can be biased.

@@

The formulation of a suitable regression model constitutes an essential part in the application of multiphase inventories. However, the requirements on the regression model properties differ fundamentally between two major mathematical approaches \cite{mandallaz2008}: In the \textit{model-dependent} (also known as \textit{model-based}) framework, the sample locations \textit{x} are fixed and the observation $Y(x)$ taken at location $x$ is assumed to be a random variable as the realization of a stochastic process (i.e the forest). As a consequence, the regression model of a model-dependent regression estimator has to met the model assumptions in the classical statistical sense, i.e. the observation of the target variable \textit{Y}(\textit{x}) is assumed to be an identically distributed, independent random variable. Moreover, the unbiasedness of the model-dependent estimators substantially depends on the correct specification of the regression model, which means that extensive analyses and adjustments must be conducted to ensure that the model adequately describes the underlying stochastic process. In contrast to the model-dependent approach, the \textit{design-based} (also known as \textit{model-assisted}) approach rests upon the randomization of the sample locations $x$ (i.e. the $x$ are independently and uniformly distributed in the forest), while the forest itself is fixed and \textit{not} the result of a stochastic process. The observation \textit{Y}(\textit{x}) itself still remains a random variable, but solemnly due to the random sample mechanism. A consequence of this approach is that the regression models used in the design-based regression estimators do no longer have to satisfy any model assumptions in the classical statistical sense (e.g. identical distribution and independence of the observation $Y(x)$, or stationarity assumption in the geostatistical frame), as it is the case in the model-based framework \citep{mandallaz2008}. The philosophy of the design-based approach is thus to use prediction models to \textit{improve} the efficiency of the estimators without having to \textit{rely} on their correct specification, which makes them very attractive to be used in official statistics. They are therefore also referred to as \textit{model-assisted}. The random distribution of the sample locations, which the design-based inference rests upon, is in practice often replaced by the use of systematic grids in order to minimize travelling cost of the terrestrial survey. However, there is reasonable and empirical evidence that softening this assumption is acceptable for point as well as variance estimation. The latter will in most cases be slightly overestimated \citep{mandallaz2013a}.

@@

\subsection{Package Structure}
\label{sec:packstruc}

In the \proglang{R} package \pkg{forestinventory}, estimators for two-phase and three-phase sampling can be applied with the \code{twophase()} and \code{threephase()} function. From these two overall function calls, various estimators for specific inventory scenarios under the chosen sampling design can be applied (figure \ref{fig:struct_package}). The decision what estimator to use thereby follows a tree-like structure which can serve the user as a guideline. The first decision to make is whether an estimate and its variance should be computed for the entire inventory area (\textbf{global estimators}) or only for subregions of the entire inventory area (\textbf{small area estimators}). The next decision for both global and small area estimations is whether the explanatory variables were derived \textit{exhaustively} or \textit{non-exhaustively} over the inventory area. The first case allows for computing the exact mean $\bar{\pmb{Z}}$ of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$). In the second case, the mean of each explanatory variable over $s_1$ ($s_{1,G}$) and $s_0$ ($s_{0,G}$) can only be estimated by $\hat{\bar{\pmb{Z}}}$ which adds another component of uncertainty to the respective estimators. The package offers three small area estimators for both the exhaustive and non-exhaustive setup. The \textit{extended synthetic estimator} and the \textit{small area estimator} both apply a bias correction using the model residuals within small area $G$ and can only be applied if the terrestrial sample size $n_{2,G}$ within $G$ is $\geq 2$. The resulting point estimates and variances are asymptotically design-unbiased. If $n_{2,G}=0$, the bias correction can no longer be applied and the estimates of the repsective \textit{synthetic small area estimator} are potentially biased. The package also allows for calculating estimates solely based on the terrestrial sample $s_2$ (\textit{one-phase estimation}) by the \code{onephase()} function. All estimators are also available for \textit{cluster} sampling, in which case a sample unit consists of multiple, spatially agglomerated sample points.

@@

The following sections describe the mathematical details and the application of the multi-phase estimators implemented in the \proglang{R} package \pkg{forestinventory}. While \cite{mandallaz2008, mandallaz2013techa, mandallaz2013techb,mandallaz2015tech} provides an extensive derivation of all estimators, we will provide the mathematical formulas that are actually implemented in the package. We will also restrict to simple sampling and provide the equations for cluster sampling in the Appendix. A special case under cluster sampling is described in section \ref{sec:speccas_and_scen}.

### 2017-08-30 rnw/Special Cases and Scenarios.Rnw
update Andreas 30.08.2017(2)

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ in included in the small area of interest. In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured.

@@

Based on the first nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e. terrestrially surveyed) and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model. If e.g. an $s_2$ and /or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:

@@

Note that that all the automatic data adjustments (deletion, recoding) have to be accepted with caution. Recapitulating, the unbiasedness of estimators in the design-based framework is based on the randomization of the sample locations. This means that every possible location within the forest area $F$ has a inclusion probability greater zero. Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures. If any information should be missing at the sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.

>>>>

For small area applications, the scenario might occur where the points of a cluster at sample locations $x$ spread over more than one small area, i.e. only a subset $M_{G}(x) < M(x)$ is included in the small area of interest. In this case, the zero mean residual property within the small area $\int_{G}M(x)R_{c}(x)d(x)=0$ is no longer insured when using the \textbf{(pseudo) extended synthetic estimator} (\textit{extsynth} and \textit{extpsynth}) which extends the set of explanatory variables by an indicator variable for the small area (see section \ref{sec:twophase_sae} and \ref{sec:threephase_sae}). In this case, one should additionally calculate the \textbf{(pseudo) small area estimator} (\textit{psmall}) where the zero mean residual property is still insured.

@@

Based on the nesting rule $s_2 \in s_1 \in s_0$, each $s_2$ (i.e. terrestrially surveyed) and $s_1$ sample location \textit{must} have all explanatory variables available that are used in the full (and thus reduced) regression model. If e.g. an $s_2$ and /or $s_1$ point misses a variable which is used in the full \textit{\textbf{and}} reduced model (here: \code{mean}), the function will \textbf{delete} this sample point from the dataset and produce the following messages:

@@

Note that that all the automatic data adjustments (deletion, recoding) have to be accepted with caution. Recapitulating, the unbiasedness of estimators in the design-based framework is based on the randomization of the sample locations. This means that every possible location within the forest area $F$ has an inclusion probability greater zero. Whereas this is already violated in practice by the use of regular grids, one can still expect that these grids do not exclude specific forest structures. If any information should be missing at the sample locations, one should clearify the reason for this and make sure that the information is \textit{completely missing at random}.

### 2017-08-30 rnw/Twophase_Estimators and their Application.Rnw
update Andreas 30.08.2017(2)

The vector regression coefficients of the OLS regression model are found by solving the sample-based normal equation.

@@

This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables. Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations. This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase} and \code{threephase} function as input data. \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g. from previous studies. However, a respective function in the package is up to now non existing and will be subject to future extentions.\par

@@

The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables. Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information. Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios. In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$. The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area, as shown in figure \ref{fig:exh_nexh_bw}. In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}). This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$. In the \textit{non-exhaustive} estimator case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}). Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}). As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par

An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase} and \code{theephase} function of the package. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g. by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}). This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy. To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}. The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.

@@

The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively. Note that this form results under the particular case where the regression coefficients are derived using the data from the current inventory (i.e. \textbf{internal} model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textbf{external} models \citep{mandallaz2016}.

@@

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance neglects the uncertainty in the regression coefficients and is thus usually slightly lower than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

@@

The structure of the data frame has to meet certain requirements to be used by the \code{twophase} and \code{threephase} functions of the package. Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location. For global two-phase estimation, we have to specify

@@

\begin{small}
<<>>=
reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
                   data=grisons,
                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
                   boundary_weights = "boundary_weights")
@
\end{small}
The \code{twophase}-function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}. A \textbf{summary of the estimation results} can be obtained by applying the \code{summary}-function to that object. It first provides a summary of the function call, the applied regression formula and the estimation method that was applied. The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}). Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model (\code{r.squared}).
\begin{small}
<<>>=
summary(reg2p_nex)
@
\end{small}
If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}. The vector must be input in the same order that the \code{lm}-function processes a \code{formula} object including the intercept term whose exact mean will always be 1. We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.

@@

We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
\begin{small}
<<>>=
reg2p_ex$estimation
@
\end{small}
Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is because we took out one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.

>>>>

The vector of regression coefficients of the OLS regression model is found by solving the sample-based normal equation

@@

This procedure is identical under both the \textit{exhaustive} and \textit{non-exhaustive} derivation of the explanatory variables. Based on the regression coefficients, the prediction for the sample location $x$ is calculated as $\pmb{Z}^t(x)\hat{\pmb{\beta}}_{s_2}$ and the empirical model residuals, which are only available at all sample locations $x \in s_2$, are calculated as $\hat{R}(x)=Y(x)-\hat{Y}(x)$. Note that in the \proglang{R} package \pkg{forestinventory}, only \textit{internal models} can be used to calculate estimations. This means that the regression model and thus the vector of regression coefficients is always derived from the current inventory data that are passed to the \code{twophase()} and \code{threephase()} function as input data. \citet{massey2015a} also describe the use of external models in the design based framework where the regression coefficients can be taken e.g. from previous studies. However, a respective function in the package is up to now non existing and will be subject to future extentions.\par

@@

The \textbf{point estimates} and their \textbf{variances} can now be calculated for the case of \textbf{exhaustively} and \textbf{non-exhaustively} derived explanatory variables. Note that exhaustively available auxiliary information does not necessarily imply exhaustive calculation of the explanatory variables from that auxiliary information. Figure \ref{fig:exh_nexh_bw} illustrates the two scenarios. In both cases, a set $\pmb{Z}$ of $p$ explanatory variables $x_{1},...,x_{p}$ is calculated within a defined spatial extent (so called \textit{support}) around each sample location $x$. The model-assisted regression estimators proposed by \citet{mandallaz2013a, mandallaz2013b} require that the mean of each explanatory variable over the entire inventory area can theoretically be computed, which is only possible if the spatial arrangement of the supports can provide a perfect tesselation of the inventory area (figure \ref{fig:exh_nexh_bw}). In the \textit{exhaustive} estimator case, the explanatory variables are indeed calculated at \textit{every} sample location $x$ over the inventory area (figure \ref{fig:exh_nexh_and_boundweights_a}). This allows for calculating the \textit{exact} mean $\bar{\pmb{Z}}=\frac{1}{n_1}\sum_{x \in s_1}\pmb{Z}(x)$. In the \textit{non-exhaustive} case, the explanatory variables are only computed at a subset of all theoretically possible sample locations (figure \ref{fig:exh_nexh_and_boundweights_b}). Consequently, the true mean $\bar{\pmb{Z}}$ has to be replaced by the \textit{estimated} mean $\hat{\bar{\pmb{Z}}}$, which adds another component of uncertainty to the estimators (compare equation \ref{eq:gw_var_2p_reg_nex} with \ref{eq:gw_var_2p_reg_ex}). As is obvious, a perfect tesselation by the supports strongly depends on the distance between the sample locations of $s_1$ as well as the support sizes and might not always be feasible in practice, in which case the assumption of tesselation must be relaxed.\par

An extension to the so-far published estimators by Mandallaz is the consideration of a \textit{\textbf{boundary adjustment}} according to forest/non-forest decision on the support level, which can be optionally applied in the \code{twophase()} and \code{theephase()} function of the package. In forest inventories, the sample is often restricted to those sample locations located within the forest area. In case a consistent forest definition can be applied to both the $s_2$ and $s_1$ sample (e.g. by a ploygon forest mask layer), it might be desired to restrict the calculation of the explanatory variables to the forest area within the given support (figure \ref{fig:exh_nexh_bw}). This method was suggested by \citet{mandallaz2013b} and led to an improvement of the regression model accuracy. To  insure  an  unbiased  calculation  of  the mean of $\pmb{Z}$, the  respective mean have then to be calculated as the \textit{weighted} mean according to equation \ref{eq:wmeanZ}. The weight $w(x)$ is thereby equal to the percentage of forested area within the support of sample location $x$.

@@

The \textbf{point estimate} for exhaustive ($ex$) and non-exhaustive ($nex$) explanatory variables is then calculated according to equation \ref{eq:pointest_2p_reg_ex} and  \ref{eq:pointest_2p_reg_nex} respectively. Note that this form results under the particular case where the regression coefficients are derived using the data from the current inventory (i.e. \textit{internal} model). In this case, the mean of the residuals $\frac{1}{n_2}\sum_{x\in{s_2}}\hat{R}(x)$ is zero by construction and does not have to be added as is necessary for \textit{external} models \citep{mandallaz2016}.

@@

The package \pkg{forestinventory} additionally provides the \textbf{external variance} (equation \ref{eq:varexternal_2p_reg_ex} and \ref{eq:varexternal_2p_reg_nex}). Note that the external variance assumes a true external regression model and thus neglects the uncertainty in the regression coefficients. It is usually slightly smaller than the design-based variance, where this uncertainty is considered by the variance-covariance matrix of the regression coefficients.

@@

The structure of the data frame has to meet certain requirements to be usable by the \code{onephase}, \code{twophase()} and \code{threephase()} function of the package. Firstly, a row in the data frame corresponds to a sample locations while the columns specify the attributes asociated to the respective sample location. For global two-phase estimation, we have to specify

@@

\begin{small}
<<>>=
reg2p_nex <- twophase(formula=tvol ~ mean + stddev + max + q75,
                   data=grisons,
                   phase_id=list(phase.col = "phase_id_2p", terrgrid.id = 2),
                   boundary_weights = "boundary_weights")
@
\end{small}
The \code{twophase()} function creates an \code{S3} object of class \code{"twophase"} with subclass \code{"global"}. A \textbf{summary of the estimation results} can be obtained by passing this object to the \code{S3} generic \code{summary()} function. It first provides a summary of the function call, the applied regression formula and the estimation method that was applied. The estimation results comprise the point estimate (\code{estimate}) as well as the design-based and external variance (\code{g_variance} and \code{ext_variance}). Further given are the sample sizes of the $s_1$ and $s_2$ sample (\code{n1}, \code{n2}) as well as the R$^2$ of the regression model (\code{r.squared}).
\begin{small}
<<>>=
summary(reg2p_nex)
@
\end{small}
If the exact means of the explanatory variables $\bar{\pmb{Z}}$ are available, the \textbf{exhaustive estimator} can be applied by additionally passing a vector containg the exact means to the optional argument \code{exhaustive}. The vector must be input in the same order that the \code{lm()} function processes a \code{formula} object including the intercept term whose exact mean will always be 1. We assume that if desired, the boundary adjustment has already been applied when deriving the exact mean of $\pmb{Z}$.

@@

We can have a look on the estimation results by particularly querying the \code{estimation} object of \code{reg2p_ex}:
\begin{small}
<<>>=
reg2p_ex$estimation
@
\end{small}
Note that both variances of the exhaustive estimation are smaller than those of the non-exhaustive estimation. This is because we eliminated one component of uncertainty by substituting the estimated means of the explanatory variables $\hat{\bar{\pmb{Z}}}$ by their exact means $\bar{\pmb{Z}}$.
